Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-19 paper.tex
save

\textit{Supervised classification.} The second part of the algorithm is the
automatic construction of a pair-wise linkage function between signatures, for
later use during clustering to group all signatures from a same author.
Formally, the goal is to build a function $d: {\cal S} \times {\cal S} \mapsto
[0, 1]$, such that $d(s_1, s_2)$ tends to $0$ if both signatures $s_1$ and
$s_2$ belong to the same author, and $1$ otherwise.

As in previous works on author disambiguation, we cast this problem as a
classification task, where inputs are pairs of signatures and outputs are
classes $0$ (same authors) and $1$ (distinct authors).

\textit{Input features.} In most cases, supervised learning algorithms assume
the input space ${\cal X}$ to be numeric (e.g., $\mathbb{R}^p$), making them
not directly applicable to structured input spaces such as ${\cal S} \times
{\cal S}$. Following previous works, pairs of signatures $(s_1, s_2)$ are
transformed to vectors $v \in \mathbb{R}^p$ by building so-called similarity
profiles \citep{treeratpituk2009disambiguating} on which supervised learning is
then carried on. In this work, we design and evaluate 15 standard input features based
on the comparison of signature fields, as reported in Table~\ref{table:features}.
As an illustrative example, the \textit{Full name} feature corresponds to ...

\begin{table}
\caption{Input features}
\label{table:features}
\centering
\begin{tabular}{|l|l|}
  \hline
  \textbf{Feature} & \textbf{Comparator}\\
  \hline
  \hline
  Full name & Cosine similarity of TF-IDF\\
  Given names & Cosine similarity of TF-IDF\\
  First given name & Jaro-Winkler distance\\
  Second given name & Jaro-Winkler distance\\
  Given name initial & Equality\\
  Affiliation & Cosine similarity of TF-IDF\\
  Co-authors & Cosine similarity of TF-IDF\\
  Title & Cosine similarity of TF-IDF\\
  Journal & Cosine similarity of TF-IDF\\
  Abstract & Cosine similarity of TF-IDF\\
  Keywords & Cosine similarity of TF-IDF\\
  Collaborations & Cosine similarity of TF-IDF\\
  References & Cosine similarity of TF-IDF\\
  Subject & Cosine similarity of TF-IDF\\
  Year difference & Absolute difference\\
  \hline
\end{tabular}
\end{table}

% ethnicity features

>>>>

\textit{Supervised classification.} The second part of the algorithm is the
automatic construction of a pair-wise linkage function between signatures, for
later use during clustering to group all signatures from a same author.
Formally, the goal is to build a function $d: {\cal S} \times {\cal S} \mapsto
[0, 1]$, such that $d(s_1, s_2)$ tends to $0$ if both signatures $s_1$ and
$s_2$ belong to the same author, and $1$ otherwise. Coincidentally, this
problem can be cast as a standard supervised classification task, where inputs
are pairs of signatures and outputs are classes $0$ (same authors) and $1$
(distinct authors).

\textit{Input features.} In most cases, supervised learning algorithms assume
the input space ${\cal X}$ to be numeric (e.g., $\mathbb{R}^p$), making them
not directly applicable to structured input spaces such as ${\cal S} \times
{\cal S}$. Following previous works \glnote{add references}, pairs of
signatures $(s_1, s_2)$ are therefore first transformed to vectors $v \in
\mathbb{R}^p$ by building so-called similarity profiles
\citep{treeratpituk2009disambiguating}, and on which supervised learning is
then carried on. In this work, we design and evaluate 15 standard input
features based on the comparison of signature fields, as reported in the first
half of Table~\ref{table:features}. As an illustrative example, the
\textit{Full name} feature corresponds to the similarity between the (full)
author name fields of the two signatures, as measured using the cosine
similarity between their respective TF-IDF vector representations as
comparator. Similarly, the \textit{Year difference} feature measures the
absolute difference between the publication date of the articles to which the
two signatures respectively belong.

Observing that author names from different cultures or origins are likely to be
disambiguated using different strategies (e.g., pairs of signatures with French
author names versus pairs of signatures with Chinese author names), and as
corroborated in \citep{treeratpituk2012name} or \citep{chin2014effective}, we
propose to complement our feature set with 7 additional new features, each
evaluating the degree of belonging of both signatures to several ethnic groups,
as reported in the second half of Table~\ref{table:features}. More
specifically, using census data extracted from \citep{rugglesintegrated}, we
built a logistic regression classifier for mapping the TF-IDF representation of
an author name to one of the 7 ethnic groups reported in that dataset. Given a
pair of signatures $(s_1, s_2)$, the proposed ethnicity features are each
computed as the estimated probability of $s_1$ to belong to the respective
ethnic group, multiplied by the estimated probability of $s_2$ to belong to the
same group. In doing so, the expectation is for the linkage function to become
sensitive to the actual origin of the authors depending on the values of these
features.

\begin{table}
\caption{Input features for learning a linkage function}
\label{table:features}
\centering
\begin{tabular}{|l|l|}
  \hline
  \textbf{Feature} & \textbf{Comparator}\\
  \hline
  \hline
  Full name & Cosine similarity of TF-IDF\\
  Given names & Cosine similarity of TF-IDF\\
  First given name & Jaro-Winkler distance\\
  Second given name & Jaro-Winkler distance\\
  Given name initial & Equality\\
  Affiliation & Cosine similarity of TF-IDF\\
  Co-authors & Cosine similarity of TF-IDF\\
  Title & Cosine similarity of TF-IDF\\
  Journal & Cosine similarity of TF-IDF\\
  Abstract & Cosine similarity of TF-IDF\\
  Keywords & Cosine similarity of TF-IDF\\
  Collaborations & Cosine similarity of TF-IDF\\
  References & Cosine similarity of TF-IDF\\
  Subject & Cosine similarity of TF-IDF\\
  Year difference & Absolute difference\\
  \hline
  White & Product of estimated probabilities\\
  Black & Product of estimated probabilities\\
  American Indian or Alaska Native & Product of estimated probabilities\\
  Chinese & Product of estimated probabilities\\
  Japanese & Product of estimated probabilities\\
  Other Asian or Pacific Islander & Product of estimated probabilities\\
  Others & Product of estimated probabilities\\
  \hline
\end{tabular}
\end{table}
